---
title: "K-means"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(formattable)
b<-colorRampPalette(brewer.pal(9,"Blues"))(100)

```

## Preparation

Load data, and split into training and testing sets
```{r}

setwd('/home/bendel/R/Hand/')
hand <-read.csv('train.csv')
train<-hand[1:30000,]
test<-hand[30001:42000,]
```

Display n-th image
```{r}
n<-1
image(matrix(as.integer(train[n,2:785]), nrow = 28, byrow=FALSE)[,c(seq(28,1))],col = b, axes = FALSE)
```


## Cluster

Use k-means to cluster into 10 groups
```{r}
set.seed(20)
clust<-kmeans(train, 10)
```

These are the centroids
```{r}
par(mfrow=c(2,5))
for (i in seq(1,10)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i, sep = ' '), col = b, axes = FALSE)
}
```

Doesn't look to promissing.

## Re-Cluster

Try a different seed

```{r}
set.seed(21)
clust<-kmeans(train, 10, iter.max = 20)
```

These are the centroids
```{r}
par(mfrow=c(2,5))
for (i in seq(1,10)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i, sep = ' '), col = b, axes = FALSE)
}
```

## Re-Cluster

Let's try using more clusters
```{r}
k <- 40
set.seed(23)
clust<-kmeans(train, k, iter.max = 30)
```

```{r, echo = FALSE}
par(mfrow=c(2,5))
for (i in seq(1,10)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i, sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(11,20)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i+10, sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(21,30)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i+20, sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(31,40)){
  image(matrix(clust$centers[i,2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', i+30, sep = ' '), col = b, axes = FALSE)
}


```

Let's try to group these. Each image comes with a label of which digit it is. The kmeans function returns the mean of the labels of all the images belonging to each cluster. Order clusters based on that value. 
```{r, echo = FALSE}
orderedCentroids <- order(clust$centers[,1])

par(mfrow=c(2,5))
for (i in seq(1,10)){
  image(matrix(clust$centers[orderedCentroids[i],2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', orderedCentroids[i], sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(11,20)){
  image(matrix(clust$centers[orderedCentroids[i],2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', orderedCentroids[i], sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(21,30)){
  image(matrix(clust$centers[orderedCentroids[i],2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', orderedCentroids[i], sep = ' '), col = b, axes = FALSE)
}
par(mfrow=c(2,5))
for (i in seq(31,40)){
  image(matrix(clust$centers[orderedCentroids[i],2:785], nrow = 28, byrow=FALSE)[,c(seq(28,1))], xlab = paste('Cluster number', orderedCentroids[i], sep = ' '), col = b, axes = FALSE)
}
```

Looks like we are getting somewhere, but far from perfect. Note that the last digits on the list are '8'. We note (see confusion matrix below) that there is a hig rate of mis-classification between '9' and '4'. The result of this is that the mean of the label of those mixed clusters is far away from '9', so those mixed clusters show up much earlier in the list, primarily mixed among the '6', and '7'.

Now label each cluster, based on visual apperance, as well as labels of centroids. Not so scientific.

```{r, echo = FALSE}
qb<-as.integer(names(clust$centers[orderedCentroids,1]))
qa<-c(0,0,0,0,0,1,1,1,2,2,2,3,3,3,4,3,5,5,5,6,4,6,6,6,4,6,6,9,9,7,9,9,7,7,7,7,8,9,8,8)

pairing<-cbind.data.frame(qb,qa)
names(pairing)<-c('Cluster', 'True')
clust1<-kmeans(test,clust$centers, iter.max = 20)
compare<-cbind.data.frame(seq(nrow(train) + 1, nrow(train) + nrow(test)), clust1$cluster)
names(compare)<-c('Image', 'Cluster')
compare$True<-NA

for (i in seq(1,40)){
  compare[compare$Cluster == pairing[i,1],3]<-pairing[i,2]
  }
```

## Evaluate

This is the confusion matrix. Actual along y-axis, predicted along x-axis.
```{r, echo = FALSE}
t<-table(compare$True, test[,1])
df<-data.frame()
for (i in seq(1,10)){
  df<-rbind.data.frame(df,t[i,])
}
names(df)<-c('0','1','2','3','4','5','6','7','8','9')
df$`True Value`<-seq(0,9)
df<-df[,c(11,1,2,3,4,5,6,7,8,9,10)]
formattable(df, list(area(col = c(2,3,4,5,6,7,8,9,10,11) ) ~ color_tile("white", "blue")))
```

This is the same confusion matrix, but here we show proportions. ie. What percent of '2' was recognized as '2', '3' etc..

```{r, echo = FALSE}
t<-table(compare$True, test[,1])
df<-data.frame()
for (i in seq(1,10)){
  df<-rbind.data.frame(df,round(t[i,]/sum(t[i,]), digits = 3))
}
names(df)<-c('0','1','2','3','4','5','6','7','8','9')
df$`True Value`<-seq(0,9)
df<-df[,c(11,1,2,3,4,5,6,7,8,9,10)]
formattable(df, list(area(col = c(2,3,4,5,6,7,8,9,10,11) ) ~ color_tile("white", "blue")))
```

```{r}
numCorrect <- 0
for (i in seq(2,11)){
   numCorrect <- numCorrect + df[i-1,i]
}
accuracy <-numCorrect/sum(df)
accuracy
```

## Conclusion

K-means clustering with 40 clusters produced correct results about 77% of the time.
